A decade of lost precision: Introducing a general framework for image feature matching without geometric constraints

### 2014-04-11 text/journal/paper.tex
Moved proofs to appendix

Given two images we can construct the proposal and baseline sets in six 
different ways as illustrated in Figure~\ref{fig:overview}. In the case 
of \emph{Ratio-Match} \cite{lowe2004sift} the \emph{proposal set} 
consists of all the feature points in the target image.  Similarly the 
\emph{baseline set} consists of all the feature points in the target 
image except the \emph{proposed} match. For \emph{Ratio-Match} this 
practically means that given a query feature, $f_q$, the \emph{proposed 
match} and the \emph{baseline match} are the two nearest neighbors in 
the set of feature points in the \emph{target image} and the ratio is 
decided uniquely based on the \emph{target image}. In turn 
\emph{Mirror-Match-Ext} as introduced by the authors in 
\cite{arnfred2013mirror}, works by constructing the \emph{proposal set} 
and \emph{baseline set} by combining features from both the \emph{query} 
and \emph{target image}.

@@

For some methods it is possible that the \emph{proposal match} is a 
feature from the \emph{query image}, in which case we discard the match 
as an untrue correspondence.  It also happens that we encounter 
correspondences where the uniqueness ratio is more than $1$ in which 
case we set it to $1$ to maintain a bounded \emph{uniqueness ratio}.  
Take for example the case of \emph{Self-Match}. Using this method we 
might find that the nearest neighbor of a feature in the target image is 
further from the query feature than the nearest neighbor in the query 
image. In this case the match is discarded.

\subsection{Generalized Uniqueness Ratio}

@@

For $F_{query} = \left\{f_{q_1} \ldots f_{q_n}\right\}$ and $F_{target} 
= \left\{f_{t_1} \ldots f_{t_m}\right\}$.  We can obtain $F_{proposed}$ 
and $F_{baseline}$ from these sets according to the algorithm used (see 
Figure~\ref{fig:overview}). $r = \text{r}(f_{q}, F_{proposed}, 
F_{baseline})$ is the ratio of query feature $f_{q}$ given 
$F_{proposed}$ and $F_{baseline}$.  If $f_{p}$ is the nearest neighbor 
of $f_{q}$ in $F_{proposed}$ and $f_{b}$ is the nearest neighbor of 
$f_{q}$ in $F_{baseline} \setminus \left\{f_{p}\right\}$ then the 
generalized \emph{uniqueness} ratio is:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \frac{d(f_{q}, f_{p})}{d(f_{q}, f_{b})}
\end{align*}
We let $K$ be the number of possible true correspondences \emph{query} 
and \emph{target image}. The \emph{precision} and \emph{recall} are then 
defined as:
\begin{align*}
    \textrm{Recall} &= \frac{\#\textrm{Correct Matches}}{K} \\
    \textrm{Precision} &= \frac{\#\textrm{Correct 
    Matches}}{\#\textrm{Correct Matches} + \#\textrm{Incorrect Matches}}
\end{align*}
Given $f_{nn} \in F_{target}$ defined as the nearest neighbor of any 
$f_q \in F_{query}$ then we define the set of features $F_{true}$ as 
$\left\{ f_{q} \in F_{query} \mid f_{nn} \text{ is a true correspondence 
of } f_{q} \right\}$ and $F_{false}$ as $F_{query} \setminus F_{true}$.  
We can the define the number of correct and incorrect matches as 
follows:
\begin{align*}
    \textrm{keep\_match}(f_{q}) &= \twopartdef{ 1 }{\text{r}(f_{q}, 
    F_{proposed}, F_{baseline}) <
    \tau}{0}{\textrm{otherwise}} \\
    \#\textrm{Correct Matches} &= \sum_{f_{q} \in F_{true}} 
    \textrm{keep\_match}(f_{q})\\
    %\mathbbm{1}_{\left\{ \right\}} \\
    \#\textrm{Incorrect Matches} &= \sum_{f_{q} \in F_{false}}
    \textrm{keep\_match}(f_{q})
    %\mathbbm{1}_{\left\{ r < \tau\right\}}
\end{align*}

@@

Based on these assumptions we first prove that the results of 
\emph{Self-Match} is identical to \emph{Self-Match-Ext}. To do this, we 
look at the nearest neighbor $f_{nn}$ of every query feature $f_{q}$ and 
consider the two cases of $f_{nn} \in F_{query}$ and $f_{nn} \in
F_{target}$. For $f_{nn} \in F_{target}$ the uniqueness ratio of 
\emph{Self-Match-Ext} is:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \text{r}(f_{q}, F_{query} \cup F_{target}, F_{baseline})\\
        &= \text{r}(f_{q}, F_{target}, F_{baseline})
\end{align*}

Since $F_{proposed} = F_{target}$ for \emph{Self-Match} this shows that
when the nearest neighbor is found in the target image, the two 
algorithms behave identically. For $f_{nn} \in F_{query}$ 
\emph{Self-Match} gives us the following ratio:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \text{r}(f_{q}, F_{target}, F_{query}) \\
        &\geq 1
\end{align*}

The closest neighbor is found in the query image, so $d(f_{q}, f_{nn}) < 
d(f_{q},f_{p})$ where $f_{p}$ is the nearest neighbor from 
$F_{proposal}$. \emph{Self-Match-Ext} does not return a correspondence 
found in the \emph{query image} and since $\tau \in [0,1]$, neither does 
\emph{Self-Match}. This proves that for any set of \emph{query features} 
\emph{Self-Match} returns the same matches as \emph{Self-Match-Ext}. For 
this reason both algorithms will later be referred to as just 
\emph{Self-Match}.

The above proof can easily be extended to show that \emph{Mirror-Match 
1} and \emph{Mirror-Match-Ext} returns the same correspondences too, and 
both will be referred to as \emph{Mirror Match} from here on.

@@

\begin{figure*}[t]
    photos of 3D objects taken 15 degrees apart. K = Keypoint, D = 
Descriptor\.}
    \caption{Keypoint / Descriptor combinations measured on 15 pairs of 
    photos of 3D objects taken 15 degrees apart. K = Keypoint, D = 
Descriptor.}
    \label{fig:descriptors}
\end{figure*}

@@

>>>>

Given two images we can construct the proposal and baseline sets in six 
different ways as illustrated in Figure~\ref{fig:overview}. In the case 
of \emph{Ratio-Match} \cite{lowe2004sift} the \emph{proposal set} 
consists of all the feature points in the target image.  Similarly the 
\emph{baseline set} consists of all the feature points in the target 
image except the \emph{proposed} match. For \emph{Ratio-Match} this 
practically means that given a query feature, $f_q$, the \emph{proposed 
match} and the \emph{baseline match} are the two nearest neighbors in 
the set of feature points in the \emph{target image} and the ratio is 
decided uniquely based on the \emph{target image}. In turn 
\emph{Mirror-Match-Ext} as introduced by the authors in 
\cite{arnfred2013mirror}, works by constructing the \emph{proposal set} 
and \emph{baseline set} by combining features from both the \emph{query} 
and \emph{target image}. By similar compositions of the \emph{proposal} 
and \emph{baseline set} we derive four more algorithms as illustrated in 
Figure~\ref{fig:overview}.

@@

For methods where the \emph{proposal set} includes the \emph{query 
features} it is possible that the \emph{proposal match} is a feature 
from the \emph{query image}, in which case we discard the match as an 
untrue correspondence.  It also happens that we encounter 
correspondences where the uniqueness ratio is more than $1$ in which 
case we set it to $1$ to maintain a bounded \emph{uniqueness ratio}.  
Take for example the case of \emph{Self-Match}. Using this method we 
might find that the nearest neighbor of a feature in the target image is 
further from the query feature than the nearest neighbor in the query 
image. In this case the match is discarded.

% TODO: include point about two of the algorithms being identical
%
\subsection{Generalized Uniqueness Ratio}
%

@@

The main distinguishing factor in between the algorithms in the 
framework is the final ratio between the distance of the two nearest 
neighbors of a query feature which decides if we keep or discard a 
match. We illustrate this process in Figure~\ref{fig:matching}.

We can define a calculation of this ratio common to all algorithms in 
the framework which allows us to compare the algorithms theoretically. 
The theoretical ratio, $r$ is defined as follows:

If $f_{p}$ is the nearest neighbor of $f_{q}$ in $F_{proposed}$ and 
$f_{b}$ is the nearest neighbor of $f_{q}$ in $F_{baseline} \setminus 
\left\{f_{p}\right\}$ then the generalized \emph{uniqueness} ratio is:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \frac{d(f_{q}, f_{p})}{d(f_{q}, f_{b})}
\end{align*}
We let $K$ be the number of possible true correspondences \emph{query} 
and \emph{target image}. The \emph{precision} and \emph{recall} are then 
defined as:
\begin{align*}
    \textrm{Recall} &= \frac{\#\textrm{Correct Matches}}{K} \\
    \textrm{Precision} &= \frac{\#\textrm{Correct 
    Matches}}{\#\textrm{Correct Matches} + \#\textrm{Incorrect Matches}}
\end{align*}
Given $f_{nn} \in F_{target}$ defined as the nearest neighbor of any 
$f_q \in F_{query}$ then we define the set of features $F_{true}$ as 
$\left\{ f_{q} \in F_{query} \mid f_{nn} \text{ is a true correspondence 
of } f_{q} \right\}$ and $F_{false}$ as $F_{query} \setminus F_{true}$.  
We can the define the number of correct and incorrect matches as 
follows:
\begin{align*}
    \textrm{keep\_match}(f_{q}) &= \twopartdef{ 1 }{\text{r}(f_{q}, 
    F_{proposed}, F_{baseline}) <
    \tau}{0}{\textrm{otherwise}} \\
    \#\textrm{Correct Matches} &= \sum_{f_{q} \in F_{true}} 
    \textrm{keep\_match}(f_{q})\\
    %\mathbbm{1}_{\left\{ \right\}} \\
    \#\textrm{Incorrect Matches} &= \sum_{f_{q} \in F_{false}}
    \textrm{keep\_match}(f_{q})
    %\mathbbm{1}_{\left\{ r < \tau\right\}}
\end{align*}

@@

Based on these assumptions we first prove that 
The above proof can easily be extended to show that \emph{Mirror-Match 
1} and \emph{Mirror-Match-Ext} returns the same correspondences too, and 
both will be referred to as \emph{Mirror Match} from here on.

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=2\columnwidth]{images/results_descriptors_imageset1}
    \caption{Keypoint / Descriptor combinations measured on 15 pairs of 
    photos of 3D objects taken 15 degrees apart. K = Keypoint, D = 
Descriptor.}
    \label{fig:descriptors}
\end{figure*}

@@

\appendix
\section{Proof of equivalence between \emph{Self-Match} and 
\emph{Self-Match-Ext}}
In this appendix we prove the equivalence between \emph{Self-Match} and 
\emph{Self-Match-Ext}. To do this, we look at the nearest neighbor 
$f_{nn}$ of every query feature $f_{q}$ and consider the two cases of 
$f_{nn} \in F_{query}$ and $f_{nn} \in
F_{target}$. For $f_{nn} \in F_{target}$ the uniqueness ratio of 
\emph{Self-Match-Ext} is:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \text{r}(f_{q}, F_{query} \cup F_{target}, F_{baseline})\\
        &= \text{r}(f_{q}, F_{target}, F_{baseline})
\end{align*}

Since $F_{proposed} = F_{target}$ for \emph{Self-Match} this shows that
when the nearest neighbor is found in the target image, the two 
algorithms behave identically. For $f_{nn} \in F_{query}$ we know that 
the best match for the query feature is in the query image itself. Since 
neither algorithm returns a correspondence within the same image, they 
also behave identically for this case. This proves that for any set of 
\emph{query features} \emph{Self-Match} returns the same matches as 
\emph{Self-Match-Ext}.

\section{Proof of equivalence between \emph{Mirror-Match} and 
\emph{Mirror-Match-Ext}}

In this appendix we prove the equivalence between \emph{Mirror-Match} 
and \emph{Mirror-Match-Ext}. This proof is almost identical to the 
equivalence proof between \emph{Self-Match} and \emph{Self-Match-Ext}. 
To prove this, we look at the nearest neighbor $f_{nn}$ of every query 
feature $f_{q}$ and consider the two cases of $f_{nn} \in F_{query}$ and 
$f_{nn} \in
F_{target}$. For $f_{nn} \in F_{target}$ the uniqueness ratio of 
\emph{Mirror-Match-Ext} is:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \text{r}(f_{q}, F_{query} \cup F_{target}, F_{baseline})\\
        &= \text{r}(f_{q}, F_{target}, F_{baseline})
\end{align*}

Since $F_{proposed} = F_{target}$ for \emph{Mirror-Match} this shows that
when the nearest neighbor is found in the target image, the two 
algorithms behave identically. For $f_{nn} \in F_{query}$ we know that 
the best match for the query feature is in the query image itself. Since 
neither algorithm returns a correspondence within the same image, they 
also behave identically for this case. This proves that for any set of 
\emph{query features} \emph{Mirror-Match} returns the same matches as 
\emph{Mirror-Match-Ext}.
